---
title: "Statistical model for scRNA-seq"
author: "Laura Jenniches"
date: "10/8/2020"
output:
  pdf_document: default
  html_document: default
---

#
```{r}
knitr::opts_chunk$set(fig.path=paste0(getwd(),"/../figures/"), warning=F,
                      message=F, echo=F, dev="cairo_pdf")
```

load R packages

```{r}
library(ggplot2)
library(ggpubr)
library(viridis)
library(ggrastr)
library(cowplot)

library(tidyverse)
library(data.table)

library(cmdstanr)
library(bayesplot)
library(posterior)

devtools::load_all("~/RNA-seq/proQ/RIF-seq_repo")

theme_set(theme_bw(base_size = 10))
```

set and create data and results directories, load gene meta data

```{r}
# set result directory for plots etc.
data_dir <- paste0(getwd(), "/../data/")
res_dir <- paste0(getwd(), "/../figures/")
if(!dir.exists(res_dir)) dir.create(res_dir)
cb_palette <- c(viridis(9)[1:8], "orange")

# load reads
meta_data <- read.table(paste0(getwd(), "/../annotations/Salmonella_meta.csv"), quote="\"", stringsAsFactors=F, header=T, sep="\t")
rownames(meta_data) <- meta_data$locus_tag
```

Read in example data (3 replicates of the first 50 genes, WT and ΔproQ only) and create long data.frame for Stan

```{r}
read_counts_in <- fread(paste0(data_dir, "proQ_RIF-seq_1-50.csv"), data.table=F)
colnames(read_counts_in)[1] <- "locus_tag"

# remove ProQ++ samples for example run
read_counts_in <- read_counts_in[,-grep("proQp", colnames(read_counts_in))]

stan_df <- read_counts_in[grep("^[Sn]",read_counts_in$locus_tag),] %>% pivot_longer(!locus_tag, names_to="sample", values_to="raw")
sample_info <- strsplit(stan_df$sample, "_")
stan_df$locus_tag <- stan_df$locus_tag %>% factor(levels=read_counts_in$locus_tag)

# relevel stan_df$condition to the control condition.
# Differences in decay rate will be given relative to the control.
stan_df$condition <- sample_info %>% sapply(function(x) x[1]) %>% factor  %>% relevel(ref="WT")
stan_df$time <- sample_info %>% sapply(function(x) sub("min", "", x[2])) %>% as.numeric()
stan_df$replicate <- sample_info %>% sapply(function(x) sub("R", "", x[3])) %>% as.numeric()

rownames(read_counts_in) <- read_counts_in$locus_tag
read_counts_in <- read_counts_in[,-grep("locus_tag", colnames(read_counts_in))]

stan_df$sample <- factor(stan_df$sample, levels=colnames(read_counts_in))
```

TMM Normalization with ERCC spike-ins

```{r}
proQ_ercc <- read_counts_in[grep("ERCC",rownames(read_counts_in)),]
# remove ERCC spike-ins with zero counts if any
remove_ercc <- which(apply(proQ_ercc, 1, prod)==0)
if(length(remove_ercc)!=0) proQ_ercc <- proQ_ercc[-remove_ercc,]
proQ_reads <- read_counts_in[grep("^[Sn]",rownames(read_counts_in)),]
nf_proQ <- get_norm_facs(proQ_reads, proQ_ercc, count_min=0, sample_min=0)
```

```{r}
stan_df$com_id <- meta_data$com_id[match(stan_df$locus_tag, meta_data$locus_tag)]
stan_df$N_lib <- colSums(proQ_reads)[match(stan_df$sample, colnames(proQ_reads))]
stan_df$n_f <- nf_proQ[match(stan_df$sample, colnames(proQ_reads))]
stan_df <- stan_df %>% mutate(cpm=(raw+0.5)*1e6/(N_lib*n_f+1), logcpm=log((raw+0.5)*1e6/(N_lib*n_f+1)))
stan_df_t0 <- subset(stan_df, time==0) %>% group_by(locus_tag, condition) %>% summarise(logcpm0=logcpm %>% mean)
stan_df$logcpm0 <- stan_df_t0$logcpm0[match(paste0(stan_df$locus_tag, stan_df$condition), paste0(stan_df_t0$locus_tag, stan_df_t0$condition))]
stan_df <- stan_df %>% mutate(logrel=logcpm - logcpm0)
stan_df_tmp <- stan_df %>% group_by(locus_tag, condition, time) %>% summarise(logrel_avg=logrel %>% mean)
stan_df$logrel_avg <- stan_df_tmp$logrel_avg[match(paste0(stan_df$locus_tag, stan_df$condition, stan_df$time), paste0(stan_df_tmp$locus_tag, stan_df_tmp$condition, stan_df_tmp$time))]
stan_df_tmp <- stan_df %>% group_by(condition, time, replicate) %>% summarise(lognf2=mean(logrel-logrel_avg))
stan_df$logrel2 <- stan_df$logrel - stan_df_tmp$lognf2[match(paste0(stan_df$time, stan_df$condition, stan_df$replicate), paste0(stan_df_tmp$time, stan_df$condition, stan_df_tmp$replicate))]
```

Verify that the center-mean normalization worked. Residuals of all replicates should be centered around 0.

```{r}
stan_df %>% subset(time==3) %>% ggplot(aes(logrel2-logrel_avg, col=(replicate %>% factor))) +
  geom_density()
```

Variables for Stan

```{r}
it_c = stan_df$condition %>% as.numeric
it_b = stan_df$replicate
it_s = stan_df$sample %>% as.numeric
it_g = stan_df$locus_tag %>% as.numeric
it_t = stan_df$time %>% factor(levels=c("0", "3", "6", "12", "24")) %>% as.numeric

N_g = unique(stan_df$locus_tag) %>% length
it_gc = (it_c - 1)*N_g + it_g
N_t = unique(stan_df$time) %>% length

data_list <- list(
  raw = stan_df$raw,
  t = stan_df$time,
  it_g = it_g,
  it_c = it_c,
  it_t = it_t,
  it_b = it_b,
  it_s = it_s,
  
  it_gc = it_gc,
  it_gct = (it_gc - 1)*N_t + it_t,
  it_gt = (it_g - 1)*N_t + it_t,
  
  N_con = max(it_c),
  N_s = max(it_s),
  N_b = stan_df %>% group_by(condition) %>% summarize(N_b=unique(it_b) %>% length) %>% pull(N_b),
  N_t = N_t,
  N_g = N_g,
  N_tot = nrow(stan_df),
  
  # TMM normalization factors and library sizes
  n_f = nf_proQ,
  N_lib = colSums(proQ_reads),

  # Remove batch effects by using center mean normalization (yes=1, no=0)
  batch_effects = 1,
  # Choose model:
  # 1: LNM
  # 2: LM
  # 3: PLM
  model_id = 1
)
```

Run cmdstan. Alternatively, data can be exported with stan_rdump

```{r}
stan_file <- paste0(getwd(), "/../stan_models/LNM_loo-1.0.stan")
mod <- cmdstan_model(stan_file, pedantic=F)

json_file <- tempfile(fileext = ".json")
write_stan_json(data_list, json_file)

# fit model with cmdstan
fit0 <- mod$sample(data = json_file, chains = 2, adapt_delta = 0.95, parallel_chains = 2, iter_warmup = 500, iter_sampling = 500, max_treedepth = 12)

# save fitted model
fit0$save_object(file = paste0(res_dir, "fit_LNM.RDS"))
rm(fit0)
cpu_time_LNM <- system.time(mod$sample(data = json_file, chains = 2, adapt_delta = 0.95, parallel_chains = 2, iter_warmup = 500, iter_sampling = 500, max_treedepth = 12))
```

Bayesian diagnostics

```{r}
# load fitted model
fit0 <- readRDS(paste0(res_dir, "fit_LNM.RDS"))

# use the bayesplot packages to check for divergent transitions and correlations between parameters
# Pairs plots
x <- fit0$draws()
i <- "LNM_1"
p <- mcmc_pairs(x, pars = paste0("delta_beta[", 1:10, "]"),
           off_diag_args = list(size = 1, alpha = 0.5), np=nuts_params(fit0))
print(p)
cairo_pdf(paste(res_dir, "pairs_", i, ".pdf", sep=""), width = 20, height = 20)
print(p)
dev.off()
png(paste(res_dir, "pairs_", i, ".png", sep=""), width = 1500, height = 1500)
print(p)
dev.off()
i <- "LNM_2"
p <- mcmc_pairs(x, pars = c("mu_beta", "beta_WT[1]", "delta_beta[1]", "sigma_g[1,1]"),
           off_diag_args = list(size = 1, alpha = 0.5), np=nuts_params(fit0))
print(p)
cairo_pdf(paste(res_dir, "pairs_", i, ".pdf", sep=""), width = 10, height = 6)
print(p)
dev.off()
png(paste(res_dir, "pairs_", i, ".png", sep=""), width = 800, height = 500)
print(p)
dev.off()

# trace plots to check whether the model fit has converged and whether all chains agree
p <- mcmc_trace(x, pars = c("mu_beta", "beta_WT[1]", "delta_beta[1]", "sigma_g[1,1]"),
           off_diag_args = list(size = 1, alpha = 0.5), np=nuts_params(fit0))
print(p)
cairo_pdf(paste(res_dir, "trace_", i, ".pdf", sep=""), width = 10, height = 6)
print(p)
dev.off()
png(paste(res_dir, "trace_", i, ".png", sep=""), width = 800, height = 500)
print(p)
dev.off()

i <- "LNM"
p <- mcmc_nuts_energy(nuts_params(fit0))
print(p)
cairo_pdf(paste(res_dir, "nuts_energy_", i, ".pdf", sep=""), width = 10, height = 6)
print(p)
dev.off()
png(paste(res_dir, "nuts_energy_", i, ".png", sep=""), width = 800, height = 500)
print(p)
dev.off()
```

Distribution of differences in decay rate

```{r}
delta_beta_draws <- fit0$draws(variables=c("delta_beta"), format="matrix")
delta_beta_proQ <- delta_beta_draws %>% apply(2, median)
beta_WT <- fit0$draws(variables=c("beta_WT"), format="matrix") %>% apply(2, median)
ggplot() +
  geom_histogram(aes(delta_beta_proQ/beta_WT))
diff <- which(delta_beta_proQ/beta_WT>0.05)
rownames(proQ_reads)[diff]
meta_data %>% subset(locus_tag %in% rownames(proQ_reads)[diff])
```

Plot decay rates
```{r}
list_param <- fit0$draws(format="matrix") %>% as.data.frame() %>% extract_params()
list_param$locus_tag <- rownames(proQ_reads)
list_param$com_id <- meta_data$com_id[match(list_param$locus_tag, meta_data$locus_tag)]
plot_genes(res_dir, subset(stan_df, com_id %in% c("ileS", "yaaJ")), subset(list_param, com_id %in% c("ileS", "yaaJ")), model_id="LNM", addNorm = F, addDect = F, show.int=F, col=cb_palette_4[c(1,2)])
```

Fit log-normal model with count-dependent variance

```{r}
stan_file <- paste0(getwd(), "/../stan_models/LNMcdv_loo-1.0.stan")
mod <- cmdstan_model(stan_file, pedantic=F)

json_file <- tempfile(fileext = ".json")
write_stan_json(data_list, json_file)

# fit model with cmdstan
fit_voom <- mod$sample(data = json_file, chains = 2, adapt_delta = 0.95, parallel_chains = 2, iter_warmup = 500, iter_sampling = 500, max_treedepth = 12)

# save fitted model
fit_voom$save_object(file = paste0(res_dir, "fit_LNMcdv.RDS"))
rm(fit_voom)
cpu_time_LNMvoom <- system.time(mod$sample(data = json_file, chains = 2, adapt_delta = 0.95, parallel_chains = 2, iter_warmup = 500, iter_sampling = 500, max_treedepth = 12))
```

```{r}
fit_voom <- readRDS(paste0(res_dir, "fit_LNMcdv.RDS"))
list_param_voom <- fit_voom$draws(format="matrix") %>% as.data.frame() %>% extract_params()
list_param_voom$locus_tag <- rownames(proQ_reads)
list_param_voom$com_id <- meta_data$com_id[match(list_param_voom$locus_tag, meta_data$locus_tag)]
plot_genes(res_dir, subset(stan_df, com_id %in% c("ileS", "yaaJ")), subset(list_param, com_id %in% c("ileS", "yaaJ")), model_id="LNMcdv", addNorm = F, addDect = F, show.int=F, col=cb_palette_4[c(1,2)])
```

```{r}
a <- fit_voom$summary("a", "median")$median
lograw_max <- max(log(stan_df$raw+1))
variance_trend <- function(log_raw){
  1-a*log_raw/log_raw_max
}

stan_df_stats <- stan_df %>% group_by(locus_tag, time) %>% summarize(logcpm_sd=sd(logcpm), logcpm_mean=mean(logcpm))

stan_df$logcpm_mean <- stan_df_stats$logcpm_mean[match(
  paste(stan_df$locus_tag, stan_df$time),
  paste(stan_df_stats$locus_tag, stan_df_stats$time))]
stan_df$logcpm_sd <- stan_df_stats$logcpm_sd[match(
  paste(stan_df$locus_tag, stan_df$time),
  paste(stan_df_stats$locus_tag, stan_df_stats$time))]

list_param_sigma <- list_param_voom %>% pivot_longer(matches("sigma"), names_to="time", names_prefix="sigma_g", values_to="sigma")
list_param_sigma$time <- c(0,3,6,12,24)[match(list_param_sigma$time, (1:5) %>% as.character())]
list_param_sigma$logcpm <- stan_df_stats$logcpm_mean[match(
  paste(list_param_sigma$locus_tag, list_param_sigma$time),
  paste(stan_df_stats$locus_tag, stan_df_stats$time)
)]
list_param_sigma$logcpm_sd <- stan_df_stats$logcpm_sd[match(
  paste(list_param_sigma$locus_tag, list_param_sigma$time),
  paste(stan_df_stats$locus_tag, stan_df_stats$time)
)]

list_param_sigma_org <- list_param %>% pivot_longer(matches("sigma"), names_to="time", names_prefix="sigma_g", values_to="sigma")
list_param_sigma_org$time <- c(0,3,6,12,24)[match(list_param_sigma_org$time, (1:5) %>% as.character())]
list_param_sigma$sigma_org <- list_param_sigma_org$sigma[match(
  paste(list_param_sigma$locus_tag, list_param_sigma$time),
  paste(list_param_sigma_org$locus_tag, list_param_sigma_org$time)
)]

p_sd <- ggplot(list_param_sigma, aes(logcpm)) +
  geom_point(aes(y=logcpm_sd, col="SD"), alpha=0.4) +
  geom_point(aes(y=sigma_org, col="LNM"), alpha=0.4) +
  geom_point(aes(y=sigma*(1-a*logcpm/max(logcpm)), col="LNMcdv"), alpha=0.4) +
  scale_color_manual(name="", values=c(viridis(3)[1:2], "orange", "gray")) +
  theme(legend.title=element_text(face="bold"),
        legend.text=element_text(size=8),
        legend.key.width=unit(0.4,"cm"),legend.key.height=unit(0.4,"cm"), legend.margin=margin(t=-0.2, b=-0.2, unit="cm"),
        legend.position="top")
p_sd_diff <- ggplot(list_param_sigma, aes(logcpm)) +
  geom_point(aes(y=sigma*(1-a*logcpm/max(logcpm)) - sigma_org), col=viridis(1), alpha=0.4) +
  ylab(expression(sigma[cdv] - sigma[LNM])) +
  theme(legend.title=element_text(face="bold"),
        legend.text=element_text(size=8),
        legend.key.width=unit(0.4,"cm"),legend.key.height=unit(0.4,"cm"),legend.margin=margin(t=-0.2, b=-0.2, unit="cm"),
        legend.position="top")

p <- cowplot::plot_grid(
  p_sd, p_sd_diff,
  ncol=2, labels=c('A', 'B'), label_size=14, scale=0.9)
print(p)
cairo_pdf(paste(res_dir, "variance_shrinkage.pdf", sep=""), width = 5, height = 2.3)
print(p)
dev.off()
```

Select parameters of interest

```{r}
calc_params <- function(fit, model_id) {
  beta_WT_draws <- fit$draws("beta_WT", format="data.frame") %>% select(contains("beta_WT"))
  beta_cond_draws <- beta_WT_draws + fit$draws("delta_beta", format="data.frame") %>% select(contains("delta_beta"))
  beta_WT_quantiles <- beta_WT_draws %>% apply(2, function(x) quantile(x, probs=c(0.05, 0.5, 0.95)))
  beta_WT <- beta_WT_quantiles[2,]
  betaWT_min <- beta_WT_quantiles[1,]
  betaWT_max <- beta_WT_quantiles[3,]
  beta_cond <- beta_cond_draws %>% apply(2, median)
  delta_beta <- beta_cond - beta_WT
  p_hl <- ((log(2)/beta_cond_draws) - (log(2)/beta_WT_draws)) %>% apply(2, function(x) p_value(x, 0.05))
  data.frame(model_id, beta_WT, betaWT_min, betaWT_max, delta_beta, p_hl)
}
```

Compare CPU time and parameters (hl_WT, delta_hl and p values) computed using the LNM, NB and LNMvoom

```{r}
data.frame(cpu_time=c(cpu_time_LNM[c(1,4)] %>% sum, cpu_time_LNMvoom[c(1,4)] %>% sum), model=c("LNM", "LNMcdv")) %>% fwrite(paste0(res_dir, "cpu_times.csv"), quote=F)
cpu_times <- fread(paste0(res_dir, "cpu_times.csv"), quote="\"")
p_col <- cpu_times %>%
  ggplot(aes(x=model, y=cpu_time)) +
  geom_col(fill=viridis(1), alpha=0.5, col="black") +
  ylab("CPU time [s]") +
  theme(legend.title=element_text(face="bold"),
        legend.text=element_text(size=8),
        legend.key.width=unit(0.4,"cm"),legend.key.height=unit(0.4,"cm"), legend.margin=margin(t=-0.2, b=-0.2, unit="cm"),
        legend.position="right")
print(p_col)
```

```{r}
gamma <- fit0$draws("gamma", format="data.frame") %>% select(contains("gamma")) %>% apply(2, median)
compare_models_df <- calc_params(fit0, "LNM")
compare_models_cdv <- calc_params(fit_voom, "LNMcdv")
colnames(compare_models_cdv) <- c("model_id", "beta_WT_alt", "betaWT_min_alt", "betaWT_max_alt", "delta_beta_alt", "p_hl_alt")
compare_models_df <- cbind(com_id=meta_data$com_id[match(rownames(proQ_reads), meta_data$locus_tag)], compare_models_df[-1], gamma, compare_models_cdv)
```

The model comparison requires uncommenting the runs related to CPU time

```{r}
p_WT <- compare_models_df %>%
  ggplot(aes(x=beta_WT, y=beta_WT_alt-beta_WT, col=model_id)) +
  scale_color_manual(name="model", values=c(viridis(3)[1], "orange")) +
  scale_alpha_continuous(name=expression(paste(t[1/2, WT], "/", CI["68%"]))) +
  xlab(expression(beta["LNM"])) +
  ylab(expression(beta["alt"] - beta["LNM"])) +
  geom_point() +
  geom_linerange(aes(ymin=betaWT_min_alt-beta_WT, ymax=betaWT_max_alt-beta_WT), alpha=0.5) +
  ggrepel::geom_label_repel(aes(label=com_id), size=3, data=compare_models_df %>% subset(model_id!="LNMrmt0" & abs(beta_WT_alt-beta_WT)>0.4)) +
  theme(legend.title=element_text(face="bold"),
        legend.text=element_text(size=8),
        legend.key.width=unit(0.3,"cm"),legend.key.height=unit(0.3,"cm"),legend.margin=margin(t=-0.2, b=-0.2, unit="cm"),
        legend.position="bottom", legend.box="vertical")
print(p_WT)

p_delta <- compare_models_df %>%
  ggplot(aes(x=delta_beta, y=delta_beta_alt-delta_beta, col=model_id, alpha=-log10(p_hl))) +
  scale_color_manual(name="model", values=c(viridis(3)[1], "orange")) +
  scale_alpha_continuous(name=expression(paste(p[max], "-", p))) +
  xlab(expression(paste( Delta, beta["LNM"]))) +
  ylab(expression(paste( Delta, beta["alt"], " - ", Delta, beta["LNM"]))) +
  geom_point() +
  theme(legend.title=element_text(face="bold"),
        legend.text=element_text(size=8),
        legend.key.width=unit(0.3,"cm"),legend.key.height=unit(0.3,"cm"),legend.margin=margin(t=-0.2, b=-0.2, unit="cm"),
        legend.position="bottom", legend.box="vertical")
print(p_delta)
p_p <- compare_models_df %>%
  ggplot(aes(x=p_hl, y=p_hl_alt-p_hl, col=model_id)) +
  scale_color_manual(name="model", values=c(viridis(3)[1], "orange")) +
  xlab(expression(p["LNM"])) +
  ylab(expression(p["alt"] - p["LNM"])) +
  geom_point() +
  theme(legend.title=element_text(face="bold"),
        legend.text=element_text(size=8),
        legend.key.width=unit(0.4,"cm"),legend.key.height=unit(0.4,"cm"),legend.margin=margin(t=-0.2, b=-0.2, unit="cm"),
        legend.position="none")
print(p_p)
p_zoom <- compare_models_df %>% subset(p_hl_alt <= 0.1 | p_hl <= 0.1) %>%
  ggplot(aes(x=p_hl, y=p_hl_alt-p_hl, col=model_id)) +
  scale_color_manual(name="model", values=c(viridis(3)[1], "orange")) +
  xlab(expression(p["LNM"])) +
  ylab(expression(p["alt"] - p["LNM"])) +
  geom_point() +
  theme(legend.title=element_text(face="bold"),
        legend.text=element_text(size=8),
        legend.key.width=unit(0.4,"cm"),legend.key.height=unit(0.4,"cm"),legend.margin=margin(t=-0.2, b=-0.2, unit="cm"),
        legend.position="none")
print(p_zoom)

p <- cowplot::plot_grid(
  cowplot::plot_grid(p_col, p_WT, p_delta, ncol=3, labels=c('A', 'B', 'C'), rel_widths=c(0.2,0.4,0.4), label_size=14, scale=0.9),
  cowplot::plot_grid(NULL, p_p, p_zoom, NULL, ncol=4, rel_widths=c(0.1,0.4,0.4,0.1), labels=c('', 'D', 'E', ''), label_size=14, scale=0.9), ncol=1, scale=1
)
print(p)
cairo_pdf(paste(res_dir, "residues_model.pdf", sep=""), width = 5.5, height = 4)
print(p)
dev.off()
```

Model comparison: correlation matrices

```{r}
cormat <- compare_models_df %>% group_by(model_id) %>% summarise(cor_beta=wCorr::weightedCorr(beta_WT, beta_WT_alt, method="Pearson", weights=1/(betaWT_max-betaWT_min)), cor_delta=wCorr::weightedCorr(delta_beta, delta_beta_alt, method="Pearson", weights=1/(betaWT_max-betaWT_min)), cor_p=cor(p_hl, p_hl_alt, method="spearman"))
```

Model comparison with LOO-CV, requires log_lik to be calculated in the generated quantities section of the stan model

```{r}
loo_lnm <- fit0$loo(cores = 3)
loo_cdv <- fit_voom$loo(cores = 3)
```

Model comparison with posterior predictive checking, requires y_rep to be calculated in the generated quantities section of the stan model

```{r}
yrep_res <- fit0$draws("y_rep", format="matrix")[1:50,] %>% t %>% as.data.frame()
yrep_res$logrel <- stan_df$logrel2
yrep_res$data_id <- 1:nrow(yrep_res)
yrep_res$raw <- stan_df$raw
yrep_res_long <- yrep_res %>% subset(raw>=10) %>% pivot_longer(!c(data_id, raw, logrel), values_to="count", names_to="draw")
yrep_stats <- yrep_res_long %>% group_by(draw) %>% summarise(med=median(count-logrel, na.rm=T))
yrep_stats2 <- yrep_res_long %>% group_by(data_id, raw, logrel) %>% summarise(med=median(count, na.rm=T))
p_pred_lnm <- yrep_res_long %>% ggplot(aes(raw)) +
  geom_line(aes(y=count-logrel, group=draw, col="samples"), alpha=0.05) +
  rasterize(geom_point(aes(y=med-logrel, col="median"), alpha=0.5, data=yrep_stats2, size=0.03), dpi=150) +
  scale_color_manual(name="", values=c("green", "black")) +
  scale_y_continuous(name="Posterior predictive", limits=c(-3,3)) +
  scale_x_log10(name="raw count", limits=c(1e1,1e5)) +
  theme(axis.line=element_line(linewidth=unit(0.1, "pt")),
        legend.position="none", legend.margin=margin(t=-0.2, b=-0.2, unit="cm"),
        legend.key.width=unit(0.6,"cm"),legend.key.height=unit(0.3,"cm"),
        strip.background = element_rect(fill=NA, linewidth=0.5), panel.spacing = unit(0, "lines"),
        axis.text=element_text(color = "black")) +
  annotate("text", x=6e4, y=-2.5, label="LNM", size=3)

xplot <- yrep_res_long %>% ggplot(aes(raw)) +
  geom_density(adjust=5) +
  scale_x_log10(name="raw count", limits=c(1e1,1e5)) +
  theme(axis.line=element_line(linewidth=unit(0.1, "pt")),
        legend.position="none", legend.margin=margin(t=-0.2, b=-0.2, unit="cm"),
        legend.key.width=unit(0.6,"cm"),legend.key.height=unit(0.3,"cm"),
        strip.background = element_rect(fill=NA, linewidth=0.5), panel.spacing = unit(0, "lines"),
        panel.grid=element_blank(), axis.text=element_text(color = "black"))
xplot <- xplot + clean_theme() + rremove("legend")

yplot <- yrep_res_long %>% ggplot(aes(count-logrel)) +
  geom_line(aes(group=draw), stat="density", adjust=5, alpha=0.05, size=0.1) +
  geom_vline(aes(xintercept=med), data=yrep_stats, alpha=0.05, size=0.1) +
  scale_x_continuous(name="", limits=c(-3,3)) +
  theme(axis.line=element_line(linewidth=unit(0.1, "pt")),
        legend.position="none", legend.margin=margin(t=-0.2, b=-0.2, unit="cm"),
        legend.key.width=unit(0.6,"cm"),legend.key.height=unit(0.3,"cm"),
        strip.background = element_rect(fill=NA, linewidth=0.5), panel.spacing = unit(0, "lines"),
        panel.grid=element_blank(), axis.text=element_text(color = "black"))
yplot <- yplot + rotate() + clean_theme() + rremove("legend")
p1 <- insert_xaxis_grob(p_pred_lnm, xplot, grid::unit(.2, "null"), position = "top")
p2 <- insert_yaxis_grob(p1, yplot, grid::unit(.2, "null"), position = "right")
p_pred_lnm <- ggdraw(p2)
print(p_pred_lnm)
```

```{r}
yrep_res <- fit_voom$draws("y_rep", format="matrix")[1:50,] %>% t %>% as.data.frame()
yrep_res$data_id <- 1:nrow(yrep_res)
yrep_res$logrel <- stan_df$logrel2
yrep_res$raw <- stan_df$raw
yrep_res_long <- yrep_res %>% subset(raw>=10) %>% pivot_longer(!c(data_id, raw, logrel), values_to="count", names_to="draw")
yrep_stats <- yrep_res_long %>% group_by(draw) %>% summarise(med=median(count-logrel, na.rm=T))
yrep_stats2 <- yrep_res_long %>% group_by(data_id, raw, logrel) %>% summarise(med=median(count, na.rm=T))
p_pred_cdv <- yrep_res_long %>% ggplot(aes(raw)) +
  geom_line(aes(y=count-logrel, group=draw, col="samples"), alpha=0.05) +
  rasterize(geom_point(aes(y=med-logrel, col="median"), alpha=0.5, data=yrep_stats2, size=0.03), dpi=150) +
  scale_color_manual(name="", values=c("green", "black")) +
  scale_y_continuous(name="Posterior predictive", limits=c(-3,3)) +
  scale_x_log10(name="raw count", limits=c(1e1,1e5)) +
  theme(axis.line=element_line(linewidth=unit(0.1, "pt")),
        legend.position="none", legend.margin=margin(t=-0.2, b=-0.2, unit="cm"),
        legend.key.width=unit(0.6,"cm"),legend.key.height=unit(0.3,"cm"),
        strip.background = element_rect(fill=NA, linewidth=0.5), panel.spacing = unit(0, "lines"),
        axis.text=element_text(color = "black")) +
  annotate("text", x=3e4, y=-2.5, label="LNMcdv", size=3)

xplot <- yrep_res_long %>% ggplot(aes(raw)) +
  geom_density(adjust=5) +
  scale_x_log10(name="raw count", limits=c(1e1,1e5)) +
  theme(axis.line=element_line(linewidth=unit(0.1, "pt")),
        legend.position="none", legend.margin=margin(t=-0.2, b=-0.2, unit="cm"),
        legend.key.width=unit(0.6,"cm"),legend.key.height=unit(0.3,"cm"),
        strip.background = element_rect(fill=NA, linewidth=0.5), panel.spacing = unit(0, "lines"),
        panel.grid=element_blank(), axis.text=element_text(color = "black"))
xplot <- xplot + clean_theme() + rremove("legend")

yplot <- yrep_res_long %>% ggplot(aes(count-logrel)) +
  geom_line(aes(group=draw), stat="density", adjust=5, alpha=0.05, size=0.1) +
  geom_vline(aes(xintercept=med), data=yrep_stats, alpha=0.05, size=0.1) +
  scale_x_continuous(name="", limits=c(-3,3)) +
  theme(axis.line=element_line(linewidth=unit(0.1, "pt")),
        legend.position="none", legend.margin=margin(t=-0.2, b=-0.2, unit="cm"),
        legend.key.width=unit(0.6,"cm"),legend.key.height=unit(0.3,"cm"),
        strip.background = element_rect(fill=NA, linewidth=0.5), panel.spacing = unit(0, "lines"),
        panel.grid=element_blank(), axis.text=element_text(color = "black"))
yplot <- yplot + rotate() + clean_theme() + rremove("legend")
p1 <- insert_xaxis_grob(p_pred_cdv, xplot, grid::unit(.2, "null"), position = "top")
p2 <- insert_yaxis_grob(p1, yplot, grid::unit(.2, "null"), position = "right")
p_pred_cdv <- ggdraw(p2)
print(p_pred_cdv)
```

```{r}
p_raw <- yrep_res_long %>% ggplot(aes(raw)) +
  geom_line(aes(y=count-logrel, group=draw, col="samples"), alpha=0.05) +
  rasterize(geom_point(aes(y=med, col="median"), alpha=0.5, data=yrep_stats2, size=0.03), dpi=150) +
  scale_color_manual(name="", values=c("green", "black"))

legend <- cowplot::get_legend(
  # create some space to the left of the legend
  p_raw + theme(legend.box.margin = margin(0, 0, 0, 12), legend.position="top", legend.key.width=unit(0.3,"cm"))
)

p <- cowplot::plot_grid(
  cowplot::plot_grid(p_pred_lnm, p_pred_cdv, labels=c('A', 'B'), label_size=14, ncol=2, scale=0.95),
  legend,
  ncol=1, rel_heights=c(0.5, 0.1), scale=1)
print(p)
cairo_pdf(paste(res_dir, "posterior_predictive.pdf", sep=""), width = 6, height = 2.4)
print(p)
dev.off()
```

```{r}

loo::loo_compare(loo_lnm, loo_cdv)

stan_df$elpd_lnm <- loo_lnm$pointwise[, "elpd_loo"]
stan_df$elpd_cdv <- loo_cdv$pointwise[, "elpd_loo"]

p_loo <- stan_df %>% ggplot(aes(raw)) +
  geom_point(aes(y=elpd_lnm, col="LNM")) +
  geom_point(aes(y=elpd_cdv, col="LNMcdv")) +
  scale_color_manual(name="model", values=c(viridis(3)[1:2], "orange")) +
  scale_alpha_continuous(name=expression(paste(t[1/2, WT], "/", CI["68%"]))) +
  ylab("ELPD") +
  scale_x_log10(name="raw count") +
  theme(legend.title=element_text(face="bold"),
        legend.text=element_text(size=8),
        legend.key.width=unit(0.3,"cm"),legend.key.height=unit(0.3,"cm"),legend.margin=margin(t=-0.2, b=-0.2, unit="cm"),
        legend.position="bottom", legend.box="vertical")
print(p_loo)

p_loo <- stan_df %>% ggplot(aes(condition)) +
  geom_violin(aes(y=elpd_lnm, col="LNM")) +
  geom_violin(aes(y=elpd_cdv, col="LNMcdv")) +
  scale_color_manual(name="model", values=c(viridis(3)[1:2], "orange")) +
  scale_alpha_continuous(name=expression(paste(t[1/2, WT], "/", CI["68%"]))) +
  ylab("ELPD") +
  theme(legend.title=element_text(face="bold"),
        legend.text=element_text(size=8),
        legend.key.width=unit(0.3,"cm"),legend.key.height=unit(0.3,"cm"),legend.margin=margin(t=-0.2, b=-0.2, unit="cm"),
        legend.position="bottom", legend.box="vertical")
print(p_loo)

p_loo_hist <- stan_df %>% ggplot() +
  geom_density(aes(x=elpd_lnm, fill="LNM")) +
  geom_density(aes(x=elpd_cdv, fill="LNMcdv")) +
  scale_fill_manual(name=expression(gamma>3.5), values=c(viridis(3)[1:2], "orange")) +
  scale_alpha_continuous(name=expression(paste(t[1/2, WT], "/", CI["68%"]))) +
  xlab("ELPD") +
  theme(legend.title=element_text(face="bold"),
        legend.text=element_text(size=8),
        legend.key.width=unit(0.3,"cm"),legend.key.height=unit(0.3,"cm"),legend.margin=margin(t=-0.2, b=-0.2, unit="cm"),
        legend.position="bottom", legend.box="vertical")
print(p_loo_hist)

plot(
  loo_lnm,
  diagnostic = c("k", "n_eff"),
  label_points = FALSE,
  main = "PSIS diagnostic plot"
)
```
